Cyclone: High Availability for Persistent Key Value Stores

### 2017-12-23 paper/paper.tex
Updates.

\author{{\rm Amitabha Roy} \\ amitabha.roy@gmail.com 
	     \and {\rm Subramanya R. Dulloor} \\ dulloor@gmail.com}

@@

DPDK does not by itself provide a TCP stack. This is not a problem, since
RAFT (and indeed most consensus protocols) tolerate network losses and
reordering by design due to the need to support asynchronous communication. In
addition, most datacenter networks rarely drop or reorder packets and provide
full bisection bandwidth between servers that might serve as Cyclone replicas. We
therefore jettisoned TCP and chose to send raw IP packets encapsulated in
ethernet frames. Cyclone takes complete control of an ethernet interface,
receiving all packets directed to it, while the IP and ethernet addresses
provide sufficient information to route the packet if necessary through multiple
switches. We currently follow this communication model both for server to server
communication between replicas as well as client to server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 18 us} to as low as {\tt 5 us}, providing a significant boost
to performance.

@@

he pseudocode in Figure~\ref{fig:control_plane} describes part of the packet
handling code in Cyclone organized as event handlers triggered on receiving a
packet at the leader. We focus on only one key event for brevity: the event
where a request is received from a client. The first step is to check that this
replica is indeed the RAFT leader in the current view (term). If not, the
message is simply dropped (a timeout causes the client to try a different
server). It then prepends RAFT related information to the packet - this includes
the current term and log index for this entry. Next, it appends a pointer to the packet to 
the NVM log. Finally, it transmits the packet to follower replicas.

@@

Our solution is based on two observations. First, we \emph {do not}
require that either all mini-batches icommit to their raft logs or
none do. What we actually requires is that either all are
\emph{applied} to Rocksdb or none of them are. Put another way we
should not apply a mini-batch from a ganged operation if any of the
other mini-batches have failed to commit to their RAFT log. This means
that the client can be stateless. The second observation is that
checking whether the other minibatches have committed is easy because
the different RAFT instances can communicate through shared memory on
the same machines unlike the more general case of distributed
participants in two phase commit.

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
    \centering
    \includegraphics[scale=0.45]{results2/rocksdb.pdf}
    \caption{Rocksdb - small updates}
    \label{fig:rocksdb}
  \end{minipage}
  \begin{minipage}{.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
  \end{minipage}
  \begin{minipage}{.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
\end{minipage}
\end{figure*}

@@

\section{Conclusion}
Cyclone shows how one can leverage a small amount of non-volatile memory to
address two fundamental difficulties in adding high availability to persistent
key value stores. First, Cyclone avoids
paying the cost of the block IO interface for every update operation. This
removes block IO latency from the task of appending updates to the log as well
as sending it out over the network for replication. In turn, this enables us to
apply software packet switching techniques and concurrency to improve the
performance of replication, and making full use of available resources when
replicating log entries over the network. By optimizing the storage and network
components simultaneously, Cyclone enables high availability to persistent key
value stores without compromising on their performance.

Cyclone is available as open source software at:\\
\url{https://github.com/sdulloor/cyclone/}
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

\author{{\rm Amitabha Roy} \\ amitabha.roy@gmail.com
	     \and {\rm Subramanya R. Dulloor} \\ subramanya.r.dulloor@intel.com}

@@

DPDK does not by itself provide a TCP stack. This is not a problem,
since RAFT (and indeed most consensus protocols) tolerate network
losses and reordering by design due to the need to support
asynchronous communication. In addition, most datacenter networks
rarely drop or reorder packets and provide full bisection bandwidth
between servers that might serve as Cyclone replicas, obviating the
need for congestion control. We therefore jettisoned TCP and chose to
send raw IP packets encapsulated in ethernet frames. Cyclone takes
complete control of an ethernet interface, receiving all packets
directed to it, while the IP and ethernet addresses provide sufficient
information to route the packet if necessary through multiple
switches. We currently follow this communication model both for server
to server communication between replicas as well as client to server
communication for requests and responses. Although a detailed
evaluation is made later, switching from the kernel TCP/IP stack to
DPDK reduces the latency between machines in our testbed from {\tt 18
  us} to as low as {\tt 5 us}, providing a significant boost to
performance.

@@

The pseudocode in Figure~\ref{fig:control_plane} describes part of the
packet handling code in Cyclone organized as event handlers triggered
on receiving a packet at the leader. We focus on only one key event
for brevity: the event where a request is received from a client. The
first step is to check that this replica is indeed the RAFT leader in
the current view (term). If not, the message is simply dropped (a
timeout causes the client to try a different server). It then prepends
RAFT related information to the packet - this includes the current
term and log index for this entry. Next, it appends a pointer to the
packet to the NVM log. Finally, it transmits the packet to follower
replicas.

@@

Our solution is based on two observations. First, we \emph {do not}
require that either all mini-batches commit to their raft logs or none
do. What we actually require is that either all are \emph{applied} to
Rocksdb or none of them are. Put another way we should not apply a
mini-batch from a ganged operation if any of the other mini-batches
have failed to commit to their RAFT log. This means that the client
can be stateless, with all the decision making done at the server
side. The second observation is that checking whether the other
minibatches have committed is easy because the different RAFT
instances can communicate through shared memory on the same machines
unlike the more general case of distributed participants in two phase
commit.

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Two level: 512B payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{results2/rocksdb.pdf}
  \caption{Rocksdb - small updates}
  \label{fig:rocksdb}
\end{figure}
\begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
\end{figure}
\begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
\end{figure}

@@

\section{Conclusion}
Cyclone shows how one can leverage a small amount of non-volatile
memory to address two fundamental difficulties in adding high
availability to persistent key value stores. First, Cyclone avoids
paying the cost of the block IO interface for every update
operation. In turn, this enables us to apply software packet
switching techniques and concurrency to improve the performance of
replication, and making full use of available resources when
replicating log entries over the network. It enables deployments of
Rocksdb with native performance without any compromise on availability.
Cyclone is available as open source software at:\\
\url{https://github.com/sdulloor/cyclone/}
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
